### 2021-09-30 docs/scraps.Rmd
model update to the chainsaw model

Turbulence is fundamental to fluvial geomorphology and open-channel flows and has been the subject of extensive research across scientific and engineering disciplines for over a century. As noted in the introduction, turbulence in river flows is usually generated by shear forces at the riverbed, but also across the entire flow depth via 'form-drag' from riverbanks, meanders, bars, suspended sediment, etc. [@moogStreamReaerationNonuniform1999]. This concept is synonymous with the geomorphic idea of available stream power, or the total energy dissipated from a river reach via the liquid's potential energy while it flows downslope [@bagnoldSedimentDischargeStream1960; @bagnoldApproachSedimentTransport1966]. Both @raymondScalingGasTransfer2012a and @moogStreamReaerationNonuniform1999 confirmed that $\epsilon_D$ (simply the available stream power normalized by the water mass) is far greater than $\epsilon_S$ in open-channel flows, therefore generating the majority of fluvial TKE dissipation.

While form-drag dissipation is effectively an alternative representation of reach-scale stream power, little attention has been paid to the resulting scale-dependecy of $\epsilon_D$. Available stream power, and thus 'form-drag' shear, exists at the depth-scale and is scale-dependent: while small roughness elements near the bed produce form-drag, if they are small relative to the total flow depth they will exert little influence on near-surface dissipation [@moogStreamReaerationNonuniform1999]. This means that, if all other properties of a river flow are held constant but the channel is progressively made wider and shallower, its depth will decrease and its wetted perimeter (and therefore total frictional resistance) will increase. This concept is illustrated graphically in Figure 2 and relates the geomorphic concept of 'river channel efficiency' to turbulence: all else being constant, a less efficient channel (wide/shallow) will have greater frictional resistance and therefore form-drag (Chow, 1959). As shown via the orange arrows in Figure 2, the inefficient channel should 1) produce more turbulence, and 2) turbulence generated via form-drag should more readily reach the surface (the red arrows in Figure 2). Note that to compare wetted perimeters across many rivers with different channel areas, we use the hydraulic radius $R_h$ (channel area divided by the wetted perimeter) as compared to the mean flow depth $H$ (channel area divided by the width). An efficient channel will have a low $\frac{R_h}{H}$ ratio while an inefficient channel will have a high $\frac{R_h}{H}$ ratio that approaches 1.

![Figure 2: Conceptual model relating river channel efficiency to the turbulent dissipation rate. All else being constant, a less efficient channel shape (wide/shallow) should 1) have more form-drag and thus turbulence and 2) turbulence generated via form-drag should more readily reach the surface. This is illustrated by the orange arrows, where the inefficient channel arrows more frequently reach the surface (and influence surface turbulence- red arrows).](cache/k600_theory/conceptual_model.jpg)

This all suggests that, for incredibly inefficient river channels, $\epsilon_S \approx \epsilon_D$. The channel produces so much turbulence from form-drag, and the flow is so shallow relative to its width, that depth-scale dissipation should be nearly synonymous with near-surface dissipation and equation 2 could be equivalently written as $k\sim Sc^{-n}(v\epsilon_D)^{1/4}$. To explore when (if at all) this occurs in rivers, we fit models using 701 river $k_{600}$ measurements from @ulsethDistinctAirWater2019a. First, we defined 'channel inefficiency' using minimum $\frac{R_h}{H}$ thresholds ranging from 0.75 to 0.995 and then we fit the two models outlined in equations 3 and 4 (because all of the data are normalized to Schmidt numbers of 600, we convienantly fold the viscosity and Schmidt numbers into the model coefficients). The goal of this analysis is to find at what $\frac{R_h}{H}$ threshold does 1) equation 3's b parameter approximate 1/4 and thus mirror equation 2 and 2) equation 4's goodness-of-fit (via the coefficient of determination $r^2$) reach a suitably high value that it is deemed physically valid.
$$\mathbf{(3)} k_{600}=a(\epsilon_D)^{b}$$
$$\mathbf{(4)} k_{600}=\beta(\epsilon_D)^{1/4}$$

The results of this analysis are plotted in Figure 3a, along with channel geometry schematics to highlight the progressive shift from efficient to inefficient river channels. Equation 3's b parameter only begins to approximate 1/4 when $R_h$ is within 1% of the mean flow depth. At this point, the model's goodness-of-fit also reaches a reasonable value of `round(models_eD4[models_eD4$name == 'Rh=H',]$r2,2)`. This indicates that a river is sufficiently inefficient when its hyradulic radius approximates the mean flow depth, and at that point $\epsilon_S \approx \epsilon_D$. Figures 3b and 3c plot the model fits for this final scenario where $\frac{R_h}{H} >= 0.995$ As Figure 3c confirms, the small-eddy model is viable when a channel is so inefficient that dissipation via form-drag approximates the near-surface dissipation rate in equation 2. Figure 3b further confirms this, where the slope of the generalized model (equation 3) equals `r round(models_ulseth[models_ulseth$name == 'Rh=H',]$slope, 2)` which is nearly equivalent to 1/4.

Finally, becuase $\epsilon_D$ refers to a depth-scale quantity, we tested whether a more parsimonious model is possible by removing the 'middle-man', i.e. the Kolmogorov velocity scale that the small-eddy model is based on [@lamontEddyCellModel1970], and directly predicting *k* from depth-scale turbulence. Figure 3d plots the fitted $k=\beta_1U_*$ model. For this data, $U_*$ explains 70% of variation in $k_{600}$ and, as expected, the fitted model is nearly identical to Figure 3c in scaling dynamics and variance explained. This all indicates that depth-scale turbulent production/dissipation dominates the surface dissipation signal in extremely inefficient river channels. These findings build off of @ulsethDistinctAirWater2019a's intial result that channel slope exerts a drammatic influence on *k* scaling dynamics, further suggesting that fluvial *k* is as much a function of hydraulic geometry as the turbulence properties of the flow.