### 2021-11-16 scraps.Rmd
'manuscript_v6'

Beyond $CO_2$ emissions, *k* for various dissolved gases are of significant importance to aquatic ecologists modeling stream metabolism [e.g. @bernhardtMetabolicRegimesFlowing2018] and water quality engineers modeling river responses to pollutant loadings [@chapraSurfaceWaterQualityModeling2008]. Because of this far-reaching importance, decades of work have focused on elucidating the physical mechanisms behind *k* in oceans and lakes [@wanninkhofAdvancesQuantifyingAirSea2009 for a review] and, to a lesser extent, fluvial gas exchange [see @hallGasExchangeStreams2020 for a review].

All of these mechanistic uncertainties are then propagated through upscaling workflows when biogeochemists predict *k* across thousands of rivers [e.g @borgesGloballySignificantGreenhousegas2015; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013] most often via equation 1 coupled with hydraulic geometry (HG). HG is the scaling relationships between streamflow and river channel hydraulics [@leopoldHydraulicGeometryStream1953; @gleasonHydraulicGeometryNatural2015] that allow biogeochemists to predict river depth and velocity from streamflow models and/or streamgauges . It is currently not well understood how sensitive global estimates of fluvial gas evasion are to the specific HG model that is employed by the worker. Further, these approaches rely on either in situ discharge records or modeled streamflow which further contributes uncertainty. This is all exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, where little in situ information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

from remote sensing (RS) data. Remote sensing of river hydraulics is a burgeoning subfield within remote sensing of hydrology, often in service of remote sensing of river discharge [RSQ- @gleasonRemoteSensingRiver2020]. This is accomplished via two general approaches: 'gauged' methods which rely on in situ river data to calibrate a method to the river(s) at hand and 'ungauged' techniques which focus on hydraulic generalizability in the service of merely improving existing knowledge in data-poor domains [@gleasonRemoteSensingRiver2020]. Many, but not all, of these ungauged approaches are developed in the context of the upcoming NASA/CNES/UKSA/CSA Surface Water and Ocean Topography (SWOT) satellite mission. SWOT is expected to launch in 2022 and provide the world's first global measurements of water surface extent and elevation at novel temporal resolutions. SWOT is a wide swath radar interferometer and will sample rivers every 1 to 7 days per 21 day repeat cycle and will measure rivers wider than 100m with a goal of expanding this to rivers at least 50m wide [@biancamariaSWOTMissionIts2016]. A decade of SWOT work has explored the multi-parameter problem of estimating the river hydraulic parameters of roughness and bathymetry from remote sensing to produce the SWOT discharge product [e.g. @andreadisConstrainingAssimilationSWOT2020; @brinkerhoffConstrainingRemoteRiver2020; @durandEstimatingReachaveragedDischarge2014; @garamboisInferenceEffectiveRiver2015; @gleasonRetrievalRiverDischarge2014a; @garamboisVariationalEstimationEffective2020; @brissetAssimilationAltimetricData2018; @hagemannBAMBayesianAMHGManning2017; @larnierRiverDischargeBathymetry2020a; @oubanasDischargeEstimationUngauged2018].

-------------------------------------------------

Given that *k* can only be directly measured via a known gas concentration gradient, eddy-covariance measurements, or tracer additions to the stream [@hallGasExchangeStreams2020], SWOT data will not provide the necessary parameters to directly measure *k*. Further, these techniques are entirely infeasible at the global scale where our goal with BIKER is primarily to better understand spatiotemporal patterns of *k* rather than obtain absolute measurements of it. Therefore, we need a simple yet physically-based model for *k* to directly relate channel hydraulics to gas exchange in the large rivers that SWOT will measure. We hypothesize that this can be done by leveraging the unique geomorphology of the very wide rivers that SWOT will observe. In this section, we develop and test process-based models for *k*, selecting one that is then implemented within BIKER (section 5).

 In particular, there is uncertainty as to which models are most appropriate at different locations along the stream-to-river continuum given that river hydraulics change predictably along a river's longitudinal profile [@wangPhysicallyBasedScaling2021]. In small and steep headwaters, 'bubble-mediated gas exchange' in whitewater might lead to substantially elevated $k_{fluvial}$ where bedload is large and general channel roughness is extremely high [@mauriceInfluenceFlowBed2017; @ulsethDistinctAirWater2019].